EBASE manuscript

### 2023-01-12 manu-draft.qmd
model io table done

load(file = here('tabs/apacmptab.RData'))
```

@@

The remaining three parameters $a$ (light efficiency, $\auni$), $r$ (respiration, $\runi$), and $b$ ($\buni$) are estimated by likelihood given the observed data and prior distributions for the parameters.  The JAGS software [@Plummer03] is used with the EBASE R package [described below, @Beck22;@RCT22] to estimate the unknown parameters using Markov Chain Monte Carlo simulations in a Bayesian framework. At each time step, the change in oxygen concentration between time steps is calculated from the equation using model inputs and parameter guesses, and then a finite difference approximation is used to estimate modeled oxygen concentration. Estimations proceed using a forward in time integration where the dissolved oxygen at each time step is informed by parameter estimates from the preceding time steps.  The estimated dissolved oxygen concentration is also returned, which can be compared to observed as one measure of model performance.

@@

Table 1: EBASE parameters and outputs

Table 2: Summary stats of Fwoxy comparison

Table 3: Summary stats of Apalachicola comparison

Table 4: Key parameters form NERRS application

>>>>

load(file = here('tabs/ebaseiotab.RData'))
load(file = here('tabs/apacmptab.RData'))
```

@@

The remaining three parameters $a$ (light efficiency, $\auni$), $r$ (respiration, $\runi$), and $b$ ($\buni$) are estimated by likelihood given the observed data and prior distributions for the parameters.  The JAGS software [@Plummer03] is used with the EBASE R package [described below, @Beck22;@RCT22] to estimate the unknown parameters using Markov Chain Monte Carlo simulations in a Bayesian framework. At each time step, the change in oxygen concentration between time steps is calculated from the equation using model inputs and parameter guesses, and then a finite difference approximation is used to estimate modeled oxygen concentration. Estimations proceed using a forward in time integration where the dissolved oxygen at each time step is informed by parameter estimates from the preceding time steps.  The estimated dissolved oxygen concentration is also returned, which can be compared to observed as one measure of model performance.  All parameter inputs and outputs for EBASE are shown in `r run_reference(id = "ebaseiotab")`.

@@

```{r}
#| label: ebaseiotab
#| tbl-cap: "EBASE model parameters showing the required input, parameters derived from the model input, and model output.  The model notation shows the components decribed in @eq-ebase.  The model notation for production, respiration, and gas exchange shows both the simple parameter name and the components of @eq-ebase used to estimate the parameter."
ebaseiotab
```
